Exploring the Evolution of Energy in Radiohead's Music

### 2024-11-21 paper/paper.qmd
Add comments in the scripts

---
title: "Exploring the Evolution of Energy in Radiohead's Music"
subtitle: "Energy Levels Fluctuate Post-2000 with Positive Correlations to Valence and Loudness Across Albums"
author: 
  - Jianing Li
  - Xizi Sun
  - Yawen Tan
  - Shanjie Jiao
  - Xinqi Yue
  - Duanyi Su
thanks: "Code and data are available at: [https://github.com/YawennnnnnTan/Spotify-Data-Analysis/tree/main]"
date: today
date-format: long
abstract: "Radiohead's music shows notable variation in energy, valence, and loudness across their discography. This analysis examines the trends in energy levels over time, the distribution of energy across albums, and the relationships between energy, positivity, and loudness. The results indicate that energy levels fluctuate significantly around post-2000 and post-2015, with louder and more positive songs tending to have higher energy. Understanding these patterns provides insights into the band's evolving artistic approach and the diversity of their musical output."
format: pdf
number-sections: true
table-of-contents: true
bibliography: references.bib
---

@@

```{r}
#| include: false
#| warning: false
#| message: false
library(here)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

@@

# Create table
head_data <- head(radiohead,6)
knitr::kable(head_data[, c(10,7,12,18,36)], format = "simple",col.names = c('Energy','Album Release Year','Loudness', 'Valence','Album name'))
```

@@

# Plot the distribution of energy across albums
ggplot(radiohead, aes(x = album_name, y = energy, fill = album_name)) +
  geom_boxplot() +
  geom_point() +
  labs(x = "Album",
       y = "Energy") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")

```
@fig-album illustrates the distribution of energy levels in Radiohead's songs across different albums. Each boxplot represents the energy distribution within an album, with the box showing the median and interquartile range, while the whiskers indicate the full range of energy values. There are noticeable differences in energy levels across the albums. For example, albums like Hail to the Thief and I Might Be Wrong generally feature higher energy songs, while albums like A Moon Shaped Pool and Amnesiac exhibit lower energy levels.

@@

# Analyze the relationship between energy and valence, and add a regression line
ggplot(radiohead, aes(x = valence, y = energy)) +
  geom_point(alpha = 0.7, color = "blue") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") +  
  labs(x = "Valence (Positivity)",
       y = "Energy") +
  theme_minimal()

```
@fig-valence illustrates the relationship between valence (positivity) and energy levels in Radiohead’s songs. Each blue dot represents a song, with valence reflecting the positivity of the emotion and energy capturing the intensity of the song. The red regression line indicates a positive correlation, suggesting that as the valence increases, the energy of the song tends to rise as well. This implies that Radiohead’s songs with more positive emotions generally have higher energy.

```{r}
#| echo: false
#| warning: false
#| label: fig-loudness
#| fig-cap: "Relationship between Loudness and Energy in Radiohead's Songs"
#| 
# Plot the relationship between energy and loudness
ggplot(radiohead, aes(x = loudness, y = energy)) +
  geom_point(alpha = 0.7, color = "darkgreen") +
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  labs(x = "Loudness (dB)",
       y = "Energy") +
  theme_minimal()
```
@fig-loudness shows the relationship between energy and loudness in Radiohead's songs. The red regression line running diagonally across the plot highlights a clear positive trend: as loudness increases, energy levels tend to rise as well. This suggests that Radiohead's louder songs are generally more energetic, which aligns with the expectation that dynamic and loud compositions often carry higher energy. The spread of points around the regression line shows some variability, but the overall correlation remains strong, as evidenced by the concentration of points close to the red line.

>>>>

---
title: "Exploring the Evolution of Energy in Radiohead's Music"
subtitle: "Energy Levels Fluctuate Post-2000 with Positive Correlations to Valence and Loudness Across Albums"
author: 
  - Yawen Tan
thanks: "Code and data are available at: [https://github.com/YawennnnnnTan/Spotify-Data-Analysis/tree/main]"
date: today
date-format: long
abstract: "Radiohead's music shows notable variation in energy, valence, and loudness across their discography. This analysis examines the trends in energy levels over time, the distribution of energy across albums, and the relationships between energy, positivity, and loudness. The results indicate that energy levels fluctuate significantly around post-2000 and post-2015, with louder and more positive songs tending to have higher energy. Understanding these patterns provides insights into the band's evolving artistic approach and the diversity of their musical output."
format: pdf
number-sections: true
table-of-contents: true
bibliography: references.bib
---

@@

```{r}
#| include: false
#| warning: false
#| message: false
#Load packages
library(here)
library(tidyverse)
library(ggplot2)
library(dplyr)
```

@@

# Create table about dataset
head_data <- head(radiohead,6)
knitr::kable(head_data[, c(10,7,12,18,36)], format = "simple",col.names = c('Energy','Album Release Year','Loudness', 'Valence','Album name'))
```

@@

# Plot the distribution of energy across albums
ggplot(radiohead, aes(x = album_name, y = energy, fill = album_name)) +
  # Initialize ggplot with the 'radiohead' dataset
  # Map 'album_name' to the x-axis, 'energy' to the y-axis
  # Use 'album_name' to define the fill color for the boxplots
  geom_boxplot() +
  # Add boxplots to represent the distribution of 'energy' for each album
  geom_point() +
  # Overlay individual points for 'energy' to show spread and outliers
  labs(x = "Album", y = "Energy") +
  # Add axis labels for better readability
  theme_minimal() +
  # Use a minimalistic theme for a clean look
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
  # Rotate x-axis labels 45 degrees for better readability
  # Remove the legend as it's unnecessary for this plot
```

@fig-album illustrates the distribution of energy levels in Radiohead's songs across different albums. Each boxplot represents the energy distribution within an album, with the box showing the median and interquartile range, while the whiskers indicate the full range of energy values. There are noticeable differences in energy levels across the albums. For example, albums like Hail to the Thief and I Might Be Wrong generally feature higher energy songs, while albums like A Moon Shaped Pool and Amnesiac exhibit lower energy levels.

@@

# Analyze the relationship between energy and valence, and add a regression line
ggplot(radiohead, aes(x = valence, y = energy)) +
  # Initialize ggplot with the 'radiohead' dataset
  # Map 'valence' to the x-axis and 'energy' to the y-axis
  geom_point(alpha = 0.7, color = "blue") + 
  # Add scatter points to show the relationship between 'valence' and 'energy'
  # Set points to semi-transparent with 'alpha = 0.7' and color them blue
  geom_smooth(method = "lm", se = FALSE, color = "red") +  
  # Add a linear regression line to show the trend
  # Set 'se = FALSE' to hide the confidence interval shading
  # Use red color for the regression line
  labs(x = "Valence (Positivity)", y = "Energy") +
  # Add axis labels for 'valence' and 'energy' for clarity
  theme_minimal()
  # Apply a clean, minimalistic theme to the plot
```
@fig-valence illustrates the relationship between valence (positivity) and energy levels in Radiohead’s songs. Each blue dot represents a song, with valence reflecting the positivity of the emotion and energy capturing the intensity of the song. The red regression line indicates a positive correlation, suggesting that as the valence increases, the energy of the song tends to rise as well. This implies that Radiohead’s songs with more positive emotions generally have higher energy.

```{r}
#| echo: false
#| warning: false
#| label: fig-loudness
#| fig-cap: "Relationship between Loudness and Energy in Radiohead's Songs"
#| 
# Plot the relationship between energy and loudness
ggplot(radiohead, aes(x = loudness, y = energy)) +
  # Initialize ggplot with the 'radiohead' dataset
  # Map 'loudness' to the x-axis and 'energy' to the y-axis
  geom_point(alpha = 0.7, color = "darkgreen") +
  # Add scatter points to visualize the relationship between 'loudness' and 'energy'
  # Use a semi-transparent green color for points with 'alpha = 0.7'
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  # Add a linear regression line to show the trend
  # Set 'se = FALSE' to hide the confidence interval shading
  # Use red color for the regression line
  labs(x = "Loudness (dB)", y = "Energy") +
  # Add axis labels for 'loudness' (in decibels) and 'energy' for clarity
  theme_minimal()
  # Apply a minimalistic theme to keep the plot clean and easy to read

```
@fig-loudness shows the relationship between energy and loudness in Radiohead's songs. The red regression line running diagonally across the plot highlights a clear positive trend: as loudness increases, energy levels tend to rise as well. This suggests that Radiohead's louder songs are generally more energetic, which aligns with the expectation that dynamic and loud compositions often carry higher energy. The spread of points around the regression line shows some variability, but the overall correlation remains strong, as evidenced by the concentration of points close to the red line.
